Generality in food webs scales with species richness, not latitude

### 2014-07-07 manuscript/latitudepaper.tex
Finished rebuilding intro, working on methods.

\title{Scaling laws of trophic level and generality vary across latitude}
\author{Alyssa Cirtwill, Tamara Romanuk?}
\date{$^1$School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand}

@@

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs (networks of feeding links between species) have been used for decades to summarize the structure of ecological communities \citep{Williams2000,earlierwork}.
One of the earliest goals of constructing food webs was to measure the relationship between the species richness of
a community and its stability \citep{Dunne2006}, a goal which is still pursued today \citep{robustness studies}.
Some authors have recognized the potential for environmental effects to influence stability \citep{Shurin2007}, as well as other food web properties such as food chain length \citep{Townsend1998,Thompson2003,2004c,Young2013,Digel2014}.
Latitude is a particularly intruiging contender for an environmental influence on food web structure \citep{Shurin2007},
as latitudinal gradients in temperature \citep{}, primary productivity \citep{}, body size \citep{}, and 
species richness \citep{Schemske2009a} are among the most robust patterns in ecology[[back up from original draft]].

Latitude's strong influence on species richness, however, presents a problem for the analysis of food web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that relationships
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties that both scale with species richness and are related to candidate
explanations for the latitudinal diversity gradient.

% Explain niche space hypothesis, Schemske, Vazquez. Then make clear that niche space suggests opposite gradient to ~S^B.
%Then give hypotheses.
Here, we are particularly interested in the potential effects of latitude on scaling relationships for the number of
links per species, generality (the number of prey per predator), and vulnerability (the number of predators per prey).
We focus on these properties because they are related to one of leading candidates for a mechanism behind the 
latitudinal species-richness gradient: namely that species' niches are narrow in the tropics, reducing competition
and permitting greater adaptive radiation (reviewed in \citet{Vazquez2004}). Alternatively, it is  possible that the
higher productivity of the tropics \citep{Brown2004} may result in a broader 
niche space \citep{Davies2007} which could also allow greater diversification even if niche sizes are globally similar.
In either case, if there is less competition at lower latitudes, there should also be less generality. A higher 
proportion of non-competing specialists should also result in lower vulnerability.
Importantly, these trends (lower generality, vulnerability, and hence links per species in the tropics) are opposite
to those based on the positive scaling relationship between links per species and species richness \citep{Riede2010}.
If, therefore, there is a latitudinal effect on generality and vulnerability beyond that of the latitudinal diversity
gradient, the scaling relationships between species richness and links per species, generality, and vulnerability 
should vary over latitude.

%Among other things, a constant scaling between S and C implies that species tend to eat the same number of prey regarless of diversity \citep{Dunne2006,Pimm1982} - rephrase a bit more, this is very close.

@@

We compiled a list of 345 food webs from the University of Canberra's GlobalWeb database (www.globalwebdb.com) 
\citep{}. Of these, 302 webs had accessible original sources. 
18 source and sink webs, which describe links to or from a focal taxon, were excluded as they are likely 
to have different food web properties than webs attempting to
describe the entire community of a site \citep{Williams2002}, as were 8 plant-pollinator, 34 host-parasitoid, and 2 
competition-focused networks
\citep{Riede2010}. Of the remaining 230 food webs, we excluded two paleowebs (comprised of probable interactions 
between extinct or prehistoric species) \citep{Simenstad1978} and five ``generalized schemes'' 
\citep{Nybakken1982,Percival1929,Swan1961,Landry1977,Petipa1979,Harrison1963}, as 
we wished to measure
directly observed rather than inferred interactions. We further excluded a food web aggregated from several sites in 
the Pacific Ocean which were widely spread in both latitude and longitude and at markedly different successional stages
\citep{Vinogradov1978}, such that it is not clear which site the food web represents. [[[SI table of rejected webs, and why]]] 
[Explain Thompson, Preston, Amundsen separate from Dunne.]
This left 223 food webs with widely varying levels of taxonomic resolution, 11 of which contained humans. 
We supplemented these networks with the free-living components of eight high-resolution estuarine food webs \citep{Dunne2013,Thompson2005} and two lake food web \citep{Preston2012,Amundsen2013} that 
included parasites, as well as XX webs that have been used in previous studies of food-web scaling.

Where the latitude of a study site was not 
explicitly stated, latitude was inferred based on Google Earth \citep{} coordinates of the described site.

. We also excluded webs describing highly
modified or artificial ecosystems, and those including humans, as these could also
have different structural properties to more ``natural'' food webs.
Where several food webs were constructed from the same study site, without obvious boundaries between ``sites'' (e.g. three food
webs from Lake Nyasa), we aggregated the webs to form one summary web for the site. That is, we included all species and
interactions observed in any of the the original webs for that site in the summary web.
Similarly, we aggregated food webs created for the same location in different years. Both types of
aggregation should help to combat a known problem with food webs, namely that infrequent interactions and
rare species tend to be under-represented.

We then converted each web to a trophic-species version by aggregating species
with the same predators and prey. This helps to reduce variation in resolution
across different studies.
We then classified them both according to latitude
and ecosystem type (terrestrial, freshwater, marine, or estuary). We then calculated 17
food-web properties for each web (Table \ref{symboltable}).

In our analysis, we first addressed whether the diversity (S) or complexity (C, L/S) of our
published food webs varied over latitude (expressed as degrees away from the
equator regardless of direction) or factorial ecosystem type. We also included
year published as in independent variable to control for changing standards
for food web studies over time. We fit independent generalized linear models
for each of the logarithms of species richness, mean links per species, and
connectance. These models were used to determine any latitudinal gradients in the food-web
properties known to drive scaling relationships with other properties.

Second, we analysed the scaling relationships between the remaining 14 food-
web parameters and connectance and species richness. We used a series of
generalized linear models with a food-web parameter as the dependent variable
and latitude, ecosystem type, and their interactions with connectance and
species richness as independent variables. (parameter ~
latitude*ecotype*(log10(connectance)+log10(species richness))). For food web
properties that were proportions, we used general linear models with a
binomial error distribution. All other food web properties were log-transformed with a normal error distribution.
When modelling the standard deviation of generality, this
required the removal of 5 webs, all with 12 or fewer species, which had
standard deviations of generality of 0. These full models were then
systematically reduced using R function dredge \cite{} in package MuMIn
\cite{} in order to isolate the model with the highest AIC. These best-fitting models were 
then examined in detail. Significant slopes over species-richness or connectance were taken to indicate scaling
relationships. Significant interactions with latitude indicated that these scaling relationships varied from the tropics
to the poles. Different intercepts indicated differences between ecosystem types.

>>>>

\title{Generality in food webs scales with species richness, not latitude}
\author{Alyssa Cirtwill, Tamara Romanuk?, Daniel Stouffer?}
\date{$^1$School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand}

@@

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs (networks of feeding links between species) have been used for decades to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork}.
One of the earliest goals of constructing food webs was to measure the relationship between the species richness of
a community and its stability \citep{Dunne2006}, a goal which is still pursued today \citep{robustness studies}.
Some authors have recognized the potential for environmental effects to influence both stability \citep{Shurin2007} and 
other food web properties such as food chain length \citep{Townsend1998,Thompson2003,2004c,Young2013,Digel2014}.
Latitude is a particularly intruiging contender for an environmental influence on food web structure \citep{Shurin2007},
as latitudinal gradients in temperature \citep{}, primary productivity \citep{}, body size \citep{}, and 
species richness \citep{Schemske2009a} are among the most robust patterns in ecology[[back up from original draft]].

Latitude's strong influence on species richness, however, presents a problem for the analysis of food web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that relationships
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties that both scale with species richness and are related to candidate
explanations for the latitudinal diversity gradient, such as measures of specialization.

% Explain niche space hypothesis, Schemske, Vazquez. Then make clear that niche space suggests opposite gradient to ~S^B.
%Then give hypotheses.
A latitudinal gradient in specialization lies behind the ``latitude-niche breadth hypothesis'' \citep{Vazquez2004}.
This hypothesis predicts that decreased seasonality in the tropics should lead to more stable populations, which in
turn should evolve smaller niches \citep{Vazquez2004}. These narrow niches should therefore allow more species to 
coexist in the tropics than can at higher latitudes. Alternatively, higher productivity of the tropics \citep{Brown2004}
may result in a broader niche space \citep{Davies2007} which could also allow greater diversification even if niche 
sizes are globally similar. Although the assumptions of the latitude-niche breadth hypothesis are only equivocably 
supported \citep{Vazquez2004}, it remains a compelling potential mechanism for the latitudinal gradient in species 
richness \citep{Lappalainen2006,Krasnov2008,Slove2010}.

Importantly, the expectation of the latitude niche-breadth hypothesis that higher species richness is associated with 
reduced competition is contary to the expected increase in the mean number of interactions per species (one potential 
measure of specialization) as species richness increases \citep{Dunne2006,Riede2010}. If, therefore, there is a
distinct effect of latitude on specialization then this effect should be visible in changing scaling relationships.
Here, we tested the effect of latitude on scaling relationships between species richness and mean links per species, 
mean generality (number of prey per predator), and mean vulnerability (number of predators per prey). Together, these
properties provide a measure a species' biotic specialization - that is, its specialization in terms of interactions 
with other species.

@@

\subsection*{Data Set}
We compiled a list of XX empirical food webs from an online database (GlobalWeb webs \citep{GlobalWeb}), supplemented 
with XX food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
Of the GlobalWeb webs, XX were rejected because their original source could not be found (XX), they were source or sink
webs rather than descriptions of an entire community (XX), were focused on plant-pollinator, host-parasitoid, or 
competitive interactions (XX), described inferred interactions in an extinct community (2), were ``generalized 
schemes'' rather than being based on empirical observation (5), or because it was not clear which of a variety of 
described sites the published food web represented (1). (See Table S1 for a list of rejected webs and reasons for their
exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
interactions between them \citep{Dunne2013}. For each of the XXX webs included in the final database, we recorded the
study site latitude if possible or, where no latitude was given, obtained estimated coordinates of the described
study site using Google Earth \citep{GoogleEarth}. Webs were also classified according to ecosystem type (terrestrial,
lake, stream, estuary, or marine), year of publication (to control for changing standards of data collection over time \citep{Dunne2006}), and whether or not humans were included (11 webs included humans).

We then converted each web to a trophic-species web by aggregating species with identical predator and prey sets.
Using trophic species webs helps to reduce variation in resolution across different studies and across taxa within
a study \citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}. The number of species ($S$) and links ($L$) in each 
trophic species
web were counted and used to calculate mean links per species ($Z$), generality ($G$), and vulnerability ($V$).

\subsection*{Scaling Relationships with S}

We were then able to examine the form of the scaling relationship between each property ($Z$, $G$, and $V$) and $S$.
The scaling relationship between $Z$ and $S$ has been shown to be a power law \citep{Riede2010} of the form

\begin{equation}
\label{Power}
$Z=\alpha S^{\beta} + \epsilon $ \,,
\end{equation}

which is often re-expreseed in logarithmic form

\begin{equation}
\label{Loglog}
$log(Z) = log(\alpha) + \beta log(S) + \epsilon $ \,.
\end{equation}

However, the two forms of the relationship imply different error structures: normally-distributed, additive error in 
\ref{Power} and lognormal, multiplicative error in \ref{Loglog}. As there is no intuitive reason to believe that our
data have one error distribution over another, we follow the recommendations in \citet{Xiao2011} and compared the two
model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
data was then used to test for potential effects of latitude.

Although scaling relationships between $S$, $G$, and $V$ have not been explicitly examined (but see scaling 
relationships for the standard devaitions of each property in \citet{Riede2010}), we expect that they will follow
power laws similar to that of the relationship between $S$ and $Z$. This is because the links taken into account in
calculating $G$ and $V$ are subsets of the total links included when calculating $Z$. As with $Z$, we explicitly 
compared the error distributions of models for $G$ and $V$ using both the power-law and logarithmic formulations.

\subsection*{Effect of Latitude}

Logarithm of a sum is awful:   log(a+c) = log(a) + log(1+b^[log(c)-log(a)]) But can probably feed that into R...

Once the best formulation of the relationship between each property and $S$ had been determined, we then assessed
the impact of latitude on this relationship. There are two ways that properties might be affected by latitude:
1) the slope of the scaling relationship with $S$ might vary over latitude or 2) the residuals $\epsilon$ might
vary over latitude. In order to test both possibilities, we used models of the form

\begin{equation}

$Z=\alpha S^{\beta_{0}+\beta_{1}\delta} + \gamma\delta + \epsilon $ \,,

\end{equation}

where $\delta$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction), ....

% of  the University of Canberra's GlobalWeb database (www.globalwebdb.com) 
% \citep{}. Of these, 302 webs had accessible original sources. 
% 18 source and sink webs, which describe links to or from a focal taxon, were excluded as they are likely 
% to have different food web properties than webs attempting to
% describe the entire community of a site \citep{Williams2002}, as were 8 plant-pollinator, 34 host-parasitoid, and 2 
% competition-focused networks
% \citep{Riede2010}. Of the remaining 230 food webs, we excluded two paleowebs (comprised of probable interactions 
% between extinct or prehistoric species) \citep{Simenstad1978} and five ``generalized schemes'' 
% \citep{Nybakken1982,Percival1929,Swan1961,Landry1977,Petipa1979,Harrison1963}, as 
% we wished to measure
% directly observed rather than inferred interactions. We further excluded a food web aggregated from several sites in 
% the Pacific Ocean which were widely spread in both latitude and longitude and at markedly different successional stages
% \citep{Vinogradov1978}, such that it is not clear which site the food web represents. 
% This left 223 food webs with widely varying levels of taxonomic resolution, 11 of which contained humans.
